Trawl Summary

### 2021-07-28 docs/ms/ms.RMD
updating workflow and moving manuscript forward.
large restructure of repository. Manuscript started.

---
title: "Manuscript title here"
author: 
- name: James R Junker
  affiliation: Louisiana Universities Marine Consortium
output:
   word_document:
     reference_docx: docx_template.docx
   pdf_document:
     keep_tex: true
   html_document: default
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \linenumbers
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---

<!--- this is a note in Rmarkdown --->

@@

Disentangling the complexity of drivers and responses in community structure is especially pressing for marine systems. Marine systems have experienced greater changes than other habitats [@antao2020]. Yet, we currently lack linkages between ecological theory, primarily derived from terrestrial systems, and data in marine systems leaving unanswered questions about biodiversity changes in these habitats [@blowes2019]. We propose a framework to explore the nature of biodiversity change in an estuary bay community and connect this change to intrinsic and extrinsic drivers. This research will build towards a more coherent theory of biodiversity change through time [@shoemaker2020]. 
<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Here, we quantified over a decade of change in a shallow estuarine community. We

@@

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's educational outreach program [https://www.lumcon.edu/education-and-outreach]<https://lumcon.edu/education-and-outreach/>. Individual trawls were conducted using *trawl apparatus description...*

## Community change

@@

@@

>>>>

---
title: "Manuscript title here"
author: 
- name: James R Junker
  affiliation: Louisiana Universities Marine Consortium
output:
   word_document:
     reference_docx: docx_template.docx
   pdf_document:
     keep_tex: true
   html_document: default
header-includes:
- \usepackage{lineno}
- \usepackage{amsmath}
- \linenumbers
linestretch: 1
bibliography: refs.bib
link-citations: no
link-color: grey
csl: ecology.csl
---
```{r Rmarkdown setup, include = FALSE, echo=F, warning=FALSE, message=FALSE}

source(here::here("datascript.R"))

TB_trawl_data = TB_trawl_data %>%
   dplyr::filter(year %ni% c("2020","2021"))

```
<!--- this is a note in Rmarkdown --->

@@

Disentangling the complexity of drivers and responses in community structure is especially pressing for marine systems. Marine systems have experienced greater changes than other habitats [@antao2020]. Yet, we currently lack linkages between ecological theory, primarily derived from terrestrial systems, and data in marine systems leaving unanswered questions about biodiversity changes in these habitats [@blowes2019]. We propose a framework to explore the nature of biodiversity change in an estuary bay community and connect this change to intrinsic and extrinsic drivers. This research will build towards a more coherent theory of biodiversity change through time [@shoemaker2020].

Here, we quantified over a decade of change in a shallow estuarine community. We

@@

Trawl sampling was conducted as part of the Louisiana Universities Marine Consortium's (LUMCON) educational outreach program (https://www.lumcon.edu/education-and-outreach). Individual trawls were conducted using *trawl apparatus description...* and

## Community characterization

Species names were checked for typos and confirmed using the *gni_parse* function from the *taxize* package [@chamberlain2020] and further standardized to consistent taxonomic resolution for entire data series (Supplemental materials).

@@

```{r mobr_stats, echo = FALSE, warning=FALSE, message=FALSE, results='hide'}
total_richness = length(unique(TB_trawl_data$Common_name))

TB_trawl_sampling = TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
   dplyr::select(Date, year, Common_name, Abundance) %>%
   group_by(year) %>%
   dplyr::summarise(trawl_days = length(unique(Date)),
                    S = length(unique(Common_name)),
                    individuals = sum(Abundance, na.rm = TRUE))

TB_trawl_taxasite <- TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  dplyr::select(Date, Common_name, Abundance) %>%
  group_by(Date, Common_name) %>%
  summarise(Abundance = sum(Abundance)) %>%
  na.omit %>% group_by(Date) %>%
  # remove trawls with less than 10 individuals 
  dplyr::filter(sum(Abundance) >=10) %>%
  group_by(Date, Common_name) %>%
  pivot_wider(names_from = Common_name, values_from = Abundance, values_fill = list(Abundance = 0)) %>%
  ungroup() %>%
  dplyr::mutate(julian_date = julian(Date, origin = as.Date("2007-01-05")),
                biweekly = ceiling(julian_date/30),
                month = month(Date),
                year = year(Date)) %>%
  dplyr::select(julian_date, Date, year, month, biweekly, everything())

n_year = 50

# trawl_rarefaction analysis

trawl_mob_in = mobr::make_mob_in(comm = TB_trawl_taxasite %>% dplyr::select(-julian_date:-biweekly),
                           plot_attr = TB_trawl_taxasite %>% dplyr::select(julian_date:biweekly),
                           coord_names = "biweekly",
                           latlong = FALSE)

trawl_mob_stats = mobr::get_mob_stats(trawl_mob_in, group_var = 'year', index = c("N", "S", "S_asymp", "S_n", "S_PIE", "PIE", "pct_rare"), effort_samples = n_year, extrapolate = TRUE, boot_groups = TRUE, n_perm = 99)

group_stats = trawl_mob_stats$groups_stats %>% data.frame

S_n = max(group_stats$effort, na.rm = TRUE)

#

```

Over the sampling period from `r min(TB_trawl_sampling$year, na.rm = TRUE)` to `r max(TB_trawl_sampling$year, na.rm = TRUE)`, `r length(unique(TB_trawl_data$Date))` trawls-days were conducted yielding `r format(signif(sum(TB_trawl_data$Abundance, na.rm = TRUE),3), scientific = TRUE)` individuals. Among years, total effort generally increased, both in number of trawl-days and individuals counted, from a minimum effort of (`r TB_trawl_sampling %>% slice_min(trawl_days) %>% dplyr::select(year) %>% unlist`) to `r max(TB_trawl_sampling$trawl_days)` (`r TB_trawl_sampling %>% slice_max(trawl_days) %>% dplyr::select(year) %>% unlist`) days and from `r min(TB_trawl_sampling$individuals, na.rm = TRUE)` (`r TB_trawl_sampling %>% slice_min(individuals) %>% dplyr::select(year) %>% unlist`) to `r format(signif(max(TB_trawl_sampling$individuals, na.rm = TRUE),3), scientific = TRUE)` (`r TB_trawl_sampling %>% slice_max(individuals) %>% dplyr::select(year) %>% unlist`) individuals. In total, `r length(unique(TB_trawl_data$Common_name))` unique species were identified and new species were consistently observed throughout the sampling period (Figure 1). Within a calendar year, we estimated the asymptotic species richness (*S~asymp~*) which was relatively stable for the first five years (2007 -- 2012), reaching a minimum value of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_min(median) %>% dplyr::select(group)`. Estimated *S~asymp~* increased in 2013 to a series maximum of `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(median) %>% unlist %>% as.numeric %>% round(0)` (`r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(lower) %>% unlist %>% as.numeric %>% round(0)` -- `r group_stats %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(upper) %>% unlist %>% as.numeric %>% round(0)`) in `r group_stats %>% data.frame %>% dplyr::filter(index == "S_asymp") %>% slice_max(median) %>% dplyr::select(group)`. However, given the large differences in effort among years, variability in *S~asymp~* may be partly attributable to sampling effects among years. Therefore, we also calculated rarefied richness at `r S_n` individuals, *S~`r S_n`~*,  for each year. Here, effort-controlled species richness, *S~`r S_n`~*, diverged qualitatively from *S~asymp~* and was generally constant, ranging from `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_min(median) %>% dplyr::select(upper) %>% round(1)`) to `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(median) %>% round(1)` (`r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(lower) %>% round(1)` -- `r group_stats %>% dplyr::filter(effort > 100) %>% slice_max(median) %>% dplyr::select(upper) %>% round(1)`) (Figure 1).

```{r figure 1, echo = FALSE, warning = FALSE, message=FALSE, fig.align='center', fig.cap="Figure 1. Multiple measures of species richness within the trawl community between the years 2007 to 2019 show continual accumulation of species (purple points and line) through the timeseries. The estimated asymptotic species richness over a calendar year (Sasymp; median values and 95% percentile interval represented by yellow points and area), suggests increasing numbers of species within the community over time. However, rarefied species richness controlled to 900 individuals (S900; median values and 95% percentil interval represented by green line and area) suggests the perceived increase in Sasymp may represent an effect of increased sampling effort"}

TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  dplyr::filter(!duplicated(Common_name, nmax = (total_richness) )) %>%
  dplyr::mutate(Date = as.Date(Date)) %>%
  arrange(Date) %>%
  dplyr::mutate(richness = 1:n())-> TB_time_accum

TB_annual_accum <- TB_trawl_data %>%
  # exclude 2020 & 2021
  dplyr::filter(year %ni% c("2020","2021")) %>%
  mutate(pres = 1) %>% group_by(year) %>%
  filter(!duplicated(Common_name)) %>%
  summarise(richness = sum(pres)) %>%
  mutate(Date = as.Date(paste(as.character(year),"01","01", sep = "-"), format = "%Y-%m-%d", origin = "2000-01-01"))

# partition beta diversity into nestedness and turnover #
# uses `betapart` package #
TB_time_df <- TB_time_accum %>%
  dplyr::select(Date, year, richness) %>%
  mutate(data_type = 'continuous',
         r_lower = NA_real_,
         r_upper = NA_real_) %>%
  bind_rows(TB_annual_accum %>% 
              mutate(data_type = "ann_raw",
                     r_lower = NA_real_,
                     r_upper = NA_real_)) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(effort > 100) %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "rarefied")) %>%
  bind_rows(group_stats %>% 
              dplyr::filter(index == 'S_asymp') %>%
              dplyr::select(year ='group', richness = 'median', r_lower = 'lower', r_upper = 'upper') %>%
              dplyr::mutate(year = as.numeric(as.character(year)),
                            Date = as.Date(paste(year,"01","01", sep = "-"), format = "%Y-%m-%d"),
                            data_type = "ann_asymp")) %>%
  dplyr::mutate(data_type = factor(data_type, levels = c("continuous", "rarefied", "ann_raw", "ann_asymp"))) %>%
  dplyr::filter(data_type != "ann_raw")

rich_time <-
  ggplot(TB_time_df, aes(x = Date, y = richness, group = data_type)) +
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "ann_asymp"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
              fill = viridis(n = 3, alpha = c(1,1,0.1), option = 'viridis')[3],
              inherit.aes = FALSE)+
  geom_ribbon(data = TB_time_df %>%
                dplyr::filter(data_type == "rarefied"),
              aes(x = Date, ymin = r_lower, ymax= r_upper),
              color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[2],
              fill = viridis(n = 3, alpha = c(1,0.2,1), option = 'viridis')[2],
              inherit.aes = FALSE)+
  geom_segment(data = TB_time_df %>% 
                 dplyr::filter(data_type %ni% 'continuous') %>%
                 pivot_wider(Date, names_from = data_type, values_from = richness),
               aes(x = Date, xend = Date, y = ann_asymp, yend = rarefied ), 
               color = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')[3],
               inherit.aes = F) +
  geom_point(data = TB_time_df, aes(colour = data_type),size = 1.5) +
  geom_line(data = TB_time_df %>% dplyr::filter(!grepl('ann_asymp', data_type)), aes(colour = data_type),size = 1.2) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  scale_y_continuous(name = "Species Richness")+
  scale_colour_manual(labels = c("continuous", expression(italic("S"[900])),expression(italic("S"["asymp"]))), values = viridis(n = 3, alpha = c(1,1,1), option = 'viridis')) +
  # scale_alpha_manual(values = c(1,1,0.3))+
  theme(legend.position = c(0, 1), legend.justification = c(0, 1), legend.background = element_rect(fill = "transparent", colour = NA),
        legend.title = element_blank(), legend.text.align = 0,
        # axis.title.x = element_blank(), axis.text.x = element_blank(),
        legend.spacing.y = unit(-0.1, 'cm'))

rich_time

```

```{r figure 2, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center'}
group_stats %>%
  dplyr::filter(index == "S_PIE") %>%
  ggplot()+
  geom_point(aes(x = Date, y = median))
```

@@

# Acknowledgements

### 2021-07-28 trawl_summary.Rmd
large restructure of repository. Manuscript started.

---
title: 'Trawl Summary'
author: "Jim Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      out_dir <- '../docs';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'trawl_summary.html')) })
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

@@

TB_trawl_data <- readRDS("../data/TB_trawl_data.rds")

@@

```{r import environmental data, include = FALSE}

  temp_paths <- list.files(path = "../data/", "*temp_*", full.names = TRUE)
  temp_files <- lapply(temp_paths, read_csv, skip = 1,
                       col_types = cols(TS = col_datetime(format = "%m/%d/%Y %H:%M")))

  roundUP <- function(x,to=10){
    to*(x%/%to + as.logical(x%%to))
  }

  tempF_to_C <- function(F){(5/9)*(F-32)}

  temp_df <- bind_rows(temp_files) %>%
    rename(datetime = 'TS', temp_F = 'F') %>%
    mutate(yday = yday(datetime))

  do_paths <- list.files(path = "../data/", "*DO_*", full.names = TRUE)
  do_files <- lapply(do_paths, read_csv, skip = 1,
                     col_types = cols(TS = col_datetime(format = "%m/%d/%Y %H:%M")))

  do_df <- bind_rows(do_files) %>%
    rename(datetime = 'TS', do_mg_L = 'mg L') %>%
    mutate(yday = yday(datetime))

  hydro_paths <- list.files(path = "../data/", "hydro{1}.*.csv", full.names= TRUE)
  hydro_files <- lapply(hydro_paths, read_csv, col_type = cols(ArrayID = col_skip(), Year = col_skip(), JDay= col_skip(),
                                                               X4 = col_skip(), X5 = col_skip(), X6 = col_skip(), Time = col_skip(), X8 = col_skip(), `Date/Time` = col_character(),
                                                               DO = col_skip(), Value4 = col_double(), Depth = col_skip(), `Depth feet` = col_skip()))
  hydro_df <- bind_rows(hydro_files) %>%
    rename(datetime = 'Date/Time', temp_C = 'Temp', temp_F = 'Water Temp F', spCon_uS_cm = 'Value4', sal_psu = 'Sal') %>%
    mutate(datetime = parse_date_time(datetime, orders = c("%m/%d/%Y %H:%M", "%m/%d/%y %H:%M")),
           yday = yday(datetime))
env_long <- hydro_df %>% select(-yday, -temp_F) %>% bind_rows(temp_df %>% mutate(temp_C = tempF_to_C(temp_F)) %>% select(-yday, -temp_F)) %>%
  full_join(do_df %>% select(-yday)) %>%
  # full_join(temp_df %>% mutate(temp_C = tempF_to_C(temp_F))%>% select(-yday, - temp_F), all = TRUE) %>%
  pivot_longer(cols = temp_C:do_mg_L, names_to = "env_var", values_to = "value")

These data vary in coverage and are being updated:

`r knitr::include_graphics("../figures/env_facet_plot.png")`

@@

ggplot(spp_mod.df, aes(x = trawls, y = richness)) + geom_path(size = 2, colour = 'red') +
  geom_ribbon(aes(x = trawls, ymin = (richness - 1.96*richness_sd), ymax = (richness + 1.96*richness_sd)), colour = "red", fill = "red", alpha = 0.5) +
  scale_y_continuous(name = expression(Sigma~"species"), breaks = seq(0,roundUP(max(spp_mod.df$richness), to = 10), by = 10), labels = seq(0,roundUP(max(spp_mod.df$richness), to = 10), by = 10)) +
  scale_x_continuous(name = "Trawls", breaks = seq(0,roundUP(max(spp_mod.df$trawls), to = 10), by = 20)) +
  theme_minimal()

@@

ggplot(TB_trawl_taxasite_yr, aes(x = sites, y = richness, group = year, colour = year)) +
  geom_path(size = 1.5) +# geom_ribbon(aes(x = sites, ymin = (richness - 1.96*sd), ymax = (richness + 1.96*sd), fill = year), alpha = 0.2) +
  scale_y_continuous(name = expression(Sigma~"species"),breaks = seq(0, roundUP(max(TB_trawl_taxasite_yr$richness), to = 10), by = 10), labels = seq(0, roundUP(max(TB_trawl_taxasite_yr$richness), to = 10), by = 10)) +
  scale_x_continuous(name = "Trawls", breaks = seq(0, roundUP(max(TB_trawl_taxasite_yr$sites), to = 10), by = 5)) +
  scale_colour_viridis(discrete = TRUE) + 
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal()

@@

#create time-accumulation df from trawl data
TB_trawl_data %>%
  filter(!duplicated(Common_name)) %>%
  mutate(richness = 1:n(), 
         Date = as.Date(Date),
         richness = sort(richness))-> TB_time_accum

>>>>

---
title: 'Trawl Summary'
author: "Jim Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) { 
      out_dir <- './docs';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), out_dir, 'trawl_summary.html')) })
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

@@

TB_trawl_data <- readRDS(here::here("data/TB_trawl_data.rds"))

@@

<!-- # These data vary in coverage and are being updated: -->
<!-- #  -->
<!-- # `r knitr::include_graphics("./figures/env_facet_plot.png")` -->

@@

ggplot(spp_mod.df, aes(x = trawls, y = richness)) + geom_path(size = 2, colour = 'red') +
  geom_ribbon(aes(x = trawls, ymin = (richness - 1.96*richness_sd), ymax = (richness + 1.96*richness_sd)), colour = "red", fill = "red", alpha = 0.5) +
  scale_y_continuous(name = expression(Sigma~"species"), breaks = seq(0,plyr::round_any(max(spp_mod.df$richness), accuracy = 10, f = ceiling), by = 10), labels = seq(0,plyr::round_any(max(spp_mod.df$richness),10, ceiling), by = 10)) +
  scale_x_continuous(name = "Trawls", breaks = seq(0,plyr::round_any(max(spp_mod.df$trawls), 10, ceiling), by = 20)) +
  theme_minimal()

@@

ggplot(TB_trawl_taxasite_yr, aes(x = sites, y = richness, group = year, colour = year)) +
  geom_path(size = 1.5) +# geom_ribbon(aes(x = sites, ymin = (richness - 1.96*sd), ymax = (richness + 1.96*sd), fill = year), alpha = 0.2) +
  scale_y_continuous(name = expression(Sigma~"species"),breaks = seq(0, plyr::round_any(max(TB_trawl_taxasite_yr$richness), 10, ceiling), by = 10), labels = seq(0, plyr::round_any(max(TB_trawl_taxasite_yr$richness), 10, f = ceiling), by = 10)) +
  scale_x_continuous(name = "Trawls", breaks = seq(0, plyr::round_any(max(TB_trawl_taxasite_yr$sites), 10, f = ceiling), by = 5)) +
  scale_colour_viridis(discrete = TRUE) + 
  scale_fill_viridis(discrete = TRUE) +
  theme_minimal()

@@

#create time-accumulation df from trawl data
total_richness = length(unique(TB_trawl_data$Common_name))
TB_trawl_data %>%
  dplyr::filter(!duplicated(Common_name, nmax = total_richness )) %>%
  dplyr::mutate(Date = as.Date(Date)) %>%
  arrange(Date) %>%
  dplyr::mutate(richness = 1:n())-> TB_time_accum
